RDF Constraints to Validate

### 2015-04-17 paper/paper7/paper7-appendix.tex
finalized arXiv articles

\begin{document}
\renewcommand{\arraystretch}{1.3}
%
%
\title{RDF Constraint Types to Validate Metadata \\ on Highly-Complex Person-Level and Aggregated Data}
\subtitle{}

@@

\begin{abstract}
...

\keywords{RDF Validation, RDF Constraints, DDI-RDF Discovery Vocabulary, Disco, RDF Data Cube Vocabulary, Linked Data, Semantic Web}
\end{abstract}

@@

Bosch and Eckert initiated a comprehensive database\footnote{Publicly available at \url{http://purl.org/net/rdf-validation}.} on RDF validation requirements to collect case studies, use cases, and requirements \cite{BoschEckert2014}. 
It is continuously updated and used to evaluate and to compare various existing solutions for RDF constraint formulation and validation. 
Requirements are classified to provide a high-level view on different solutions and to facilitate a better understanding of the problem domain.
Bosch et al. identified in total 74 requirements to formulate RDF constraints; each of them corresponding to a constraint type. 
They recently published a technical report\footnote{Available at: \url{http://arxiv.org/abs/1501.03933}} in which they explain each requirement (constraint type) in detail and give examples for each (represented by listed constraint languages) \cite{BoschNolleAcarEckert2015}.
We state at least one Disco constraint for the majority of the constraint types.
Where appropriate constraint types are related to complementary requirements of the RDF validation requirements database.
When constraint types are mapped to DL, we do not state namespace prefixes for simplicity reasons.

\section{Validation Types}

We distinguish two validation types:
(1) \emph{Content-Driven Validation} $\mathcal{C}_{C}$ contains the set of constraints ensuring that the data is consistent with the intended syntax, semantics, and integrity of given data models.
(2) \emph{Technology-Driven Validation} $\mathcal{C}_{T}$ includes the set of constraints which can be generated automatically out of data models, such as cardinality restrictions, universal and existential quantifications, domains, and ranges.

\section{Validation of Metadata on Person-Level and Aggregated Data}

\begin{itemize}
  \item Constraint types marked with an asterisk (*) can be used as OWL 2 axioms. Thus, reasoners may be used to infer implicit triples resolving constraint violations caused when these constraints are validated.
	\item \underline{Underlined subsections} represent data model specific constraint types and do not correspond to RDF validation requirements (constraint types).
%	\item \underline{Underlined constraints} are part of content-driven validation.
%	\item \udash{Dashed underlined constraints} are part of technology-driven validation.
	\item $\mathcal{C}_{T}$ constraint types are marked with the superscript\textsuperscript{T}
	\item $\mathcal{C}_{C}$ constraint types are not marked
\end{itemize}

\subsection{\underline{Comparison}}

\subsection{\underline{Data Model Consistency}}
Is the data consistent with the intended semantics of the data model?
Such validation rules ensure the integrity of the data according to the data model.

\begin{itemize}
	\item \textbf{{\em DISCO-C-DATA-MODEL-CONSISTENCY-01}}: 
	Codes (\emph{skos:Concept}) are ordered and therefore have fixed positions in an ordered collection (\emph{skos:OrderedCollection}) within a variable representation.
	The cumulative percentage of the current code is the cumulative percentage of the previous code (\emph{disco:cumulativePercentage})
	plus the percentage value (\emph{disco:percentage}) of the current code. 
		\begin{itemize}
		\item severity level: ERROR
	\end{itemize}
	\item \textbf{{\em DISCO-C-DATA-MODEL-CONSISTENCY-02}}: 
	The cumulative percentage (\emph{disco:cumulativePercentage}) of the last code must be 100. 
		\begin{itemize}
		\item severity level: ERROR
	\end{itemize}
	\item \textbf{{\em DISCO-C-DATA-MODEL-CONSISTENCY-03}}: 
	The number of valid cases (\emph{disco:SummaryStatistics} of the type (\emph{disco:summaryStatisticType}) \emph{ddicv-sumstats:ValidCases}) 
	for a particular variable must exactly be the sum of absolute frequencies of all valid cases (\emph{disco:inValid} of \emph{skos:Concept} is true).
		\begin{itemize}
		\item severity level: ERROR
	\end{itemize}
	\item \textbf{{\em DISCO-C-DATA-MODEL-CONSISTENCY-04}}: 
	The number of invalid cases (\emph{disco:SummaryStatistics} of the type (\emph{disco:summaryStatisticType}) \emph{ddicv-sumstats:InvalidCases}) 
	for a particular variable must exactly be the sum of absolute frequencies of all invalid cases (\emph{disco:inValid} of \emph{skos:Concept} is false).
		\begin{itemize}
		\item severity level: ERROR
	\end{itemize}
	\item \textbf{{\em DISCO-C-DATA-MODEL-CONSISTENCY-05}}: 
	The total number of valid and invalid cases (\emph{disco:SummaryStatistics} of the type (\emph{disco:summaryStatisticType}) \emph{ddicv-sumstats:NumberOfCases}) 
	for a particular variable must exactly be the number of valid cases plus the number of invalid cases.
		\begin{itemize}
		\item severity level: ERROR
	\end{itemize}
	\item \textbf{{\em DISCO-C-DATA-MODEL-CONSISTENCY-06}}: 
	Some summary statistics types can only be calculated for given variable types. 
	It is not possible to compute minimum values for string variables.
		\begin{itemize}
		\item severity level: ERROR
	\end{itemize}
	\item \textbf{{\em DISCO-C-DATA-MODEL-CONSISTENCY-07}}: 
	Some summary statistics types can only be calculated for given variable types. 
	It is not possible to compute mean values for categorical variables, only for metric variables.
		\begin{itemize}
		\item severity level: ERROR
	\end{itemize}
\end{itemize}

@@

\begin{itemize}
	\item \textbf{{\em DISCO-C-LITERAL-RANGES-01:}}
{\em disco:percentage} (domain: {\em disco:CategoryStatistics}) literals must be of the datatype {\em xsd:double} whose range should be restricted to be between 0 and 100.
	\begin{itemize}
		\item severity level: ERROR
	\end{itemize}
\end{itemize}

@@

	\item \textbf{{\em DISCO-C-MATHEMATICAL-OPERATIONS-02}}:
	For a given variable, the sum of all category statistics (frequency) has to be equal to the 'total' value of summary statistics.
	\begin{itemize}
		\item severity level: ERROR
	\end{itemize}

	\item \textbf{{\em DISCO-C-MATHEMATICAL-OPERATIONS-03}}:
	For a given variable, the sum of 'valid total' and 'missing total' has to be equal to 'total'.
	\begin{itemize}
		\item severity level: ERROR
	\end{itemize}

	\item \textbf{{\em DISCO-C-MATHEMATICAL-OPERATIONS-04}}:
	For a given variable, the 'total' value for country 'All' must be equal to the sum of the 'total' value for all Countries.
	\begin{itemize}
		\item severity level: ERROR
	\end{itemize}

@@

\begin{itemize}
	\item \textbf{{\em DISCO-C-AGGREGATION-01}}: calculate the number of theoretical concepts in the thematic classification of a given study.
	\begin{itemize}
		\item severity level: INFO
	\end{itemize}
  \item \textbf{{\em DISCO-C-AGGREGATION-02}}: calculate the number of variables of a data set.
	\begin{itemize}
		\item severity level: INFO
	\end{itemize}
	\item \textbf{{\em DISCO-C-AGGREGATION-03}}: calculate the number of questions in a given questionnaire.
	\begin{itemize}
		\item severity level: INFO
	\end{itemize}
	\item \textbf{{\em DISCO-C-AGGREGATION-04}}: the number of codes of a given variable must be below a maximum value.
	\begin{itemize}
		\item severity level: INFO
	\end{itemize}
	\item \textbf{{\em DISCO-C-AGGREGATION-05}}: the number of questions of a given questionnaire must exactly be a given value.
	\begin{itemize}
		\item severity level: INFO
	\end{itemize}
	\item \textbf{{\em DISCO-C-AGGREGATION-06}}: the cumulative percentage of all codes of a given variable must be 100.
	\begin{itemize}
		\item severity level: INFO
	\end{itemize}
  \item \textbf{{\em DISCO-C-AGGREGATION-07}}: the absolute frequency of all valid codes of a given variable must be equal to a given value.
	\begin{itemize}
		\item severity level: INFO
	\end{itemize}
\end{itemize}

\subsection{\underline{Structure}}

@@

\subsection{\underline{Labeling and Documentation}}

@@

\subsection{\underline{Vocabulary}\textsuperscript{T}}
Vocabularies should not invent any new terms or use deprecated elements.

@@

\subsection{\underline{Linked Data}}

>>>>

\begin{document}
\renewcommand{\arraystretch}{1.3}
%
%
\title{RDF Constraints to Validate}
\subtitle{Rectangular Data and Metadata on Person-Level Data, Aggregated Data, Thesauri, and Statistical Classifications}

@@

\begin{abstract}
To ensure high quality of and trust in both metadata and data, their representation in RDF must satisfy certain criteria - specified in terms of RDF constraints.
From 2012 to 2015 together with other Linked Data community members and experts from the social, behavioural, and economic sciences (\emph{SBE}), we developed diverse vocabularies to represent \emph{SBE} metadata and rectangular data in RDF.
The \emph{DDI-RDF Discovery Vocabulary (Disco)} 
is designed to support the dissemination, management,
and reuse of person-level data, i.e., data about individuals, households, and businesses, collected in form of responses to studies and archived for research purposes.
The \emph{RDF Data Cube Vocabulary (Data Cube)} is a W3C recommendation for expressing \emph{data cubes}, i.e. multi-dimensional aggregate data. 
\emph{Physical Data Description (PHDD)} is a vocabulary to model data in rectangular format. 
The data could either be represented in records with character-separated values (\emph{CSV}) or fixed length. 
The \emph{Simple Knowledge Organization System (SKOS)} is a vocabulary to build knowledge organization systems such as thesauri, classification schemes, and taxonomies.
\emph{XKOS} is a SKOS extension to describe formal statistical classifications.

In this paper, we describe RDF constraints to validate metadata on person-level data (\emph{Disco}), aggregated data (\emph{Data Cube}), thesauri (\emph{SKOS}), and statistical classifications (\emph{XKOS}) and to validate rectangular data (\emph{PHDD}).
We assign RDF constraints to RDF constraint types either corresponding to RDF validation requirements or to data model specific constraint types.
This technical report is updated continuously as modifying, adding, and deleting constraints remains ongoing work.

\keywords{RDF Validation, RDF Constraints, DDI-RDF Discovery Vocabulary, RDF Data Cube Vocabulary, Thesauri, SKOS, Rectangular Data, Statistical Classifications, XKOS, Linked Data, Semantic Web}
\end{abstract}

@@

Bosch and Eckert initiated a comprehensive database\footnote{Publicly available at \url{http://purl.org/net/rdf-validation}.} on RDF validation requirements to collect case studies, use cases, and requirements \cite{BoschEckert2014}. 
It is continuously updated and used to evaluate and to compare various existing solutions for RDF constraint formulation and validation. 
Requirements are classified to provide a high-level view on different solutions and to facilitate a better understanding of the problem domain.
Bosch et al. identified in total 74 requirements to formulate RDF constraints; each of them corresponding to a constraint type. 
They recently published a technical report\footnote{Available at: \url{http://arxiv.org/abs/1501.03933}} in which they explain each requirement (constraint type) in detail and give examples for each (represented by listed constraint languages) \cite{BoschNolleAcarEckert2015}.
Where appropriate constraint types are related to complementary requirements of the RDF validation requirements database.

We distinguish two validation types:
(1) \emph{Content-Driven Validation} $\mathcal{C}_{C}$ contains the set of constraints ensuring that the data is consistent with the intended syntax, semantics, and integrity of given data models.
(2) \emph{Technology-Driven Validation} $\mathcal{C}_{T}$ includes the set of constraints which can be generated automatically out of data models, such as cardinality restrictions, universal and existential quantifications, domains, and ranges.
We determined the default \emph{severity level} (corresponds to requirement \emph{R-158}) for each constraint to indicate how serious the violation of the constraint is.
We propose an extensible metric to measure the continuum of severity levels ranging from $\mathcal{SL}_{0}$ (informational) via $\mathcal{SL}_{1}$ (warning) to $\mathcal{SL}_{2}$ (error). 
Although we provide default severity levels for each constraint, users should be able to specify severity levels of constraints they need to validate for their individual use cases, i.e., users should be able to define use case specific severity levels for constraints.

The following two sections encompass RDF constraints classified as constraint types either corresponding to RDF validation requirements (section \ref{RDF-validation-requirements-constraint-types}) or data model specific constraint types (section \ref{data-model-specific-constraint-types}).
Constraint types marked with an asterisk (*) can be used as OWL 2 axioms. Thus, reasoners may be used prior to the validation to infer implicit triples which may resolve or cause constraint violations.
$\mathcal{C}_{T}$ constraint types are marked with the superscript\textsuperscript{T}.
As $\mathcal{C}_{T}$ and $\mathcal{C}_{C}$ are disjoint sets, $\mathcal{C}_{C}$ constraint types are not explicitly marked.

%\section{RDF Constraints to Validate Metadata and Data}

\section{RDF Constraint Types Corresponding to RDF Validation Requirements}
\label{RDF-validation-requirements-constraint-types}

@@

\begin{itemize}
	\item \textbf{{\em DISCO-C-LITERAL-RANGES-01:}}
{\em disco:percentage} (domain: {\em disco:CategoryStatistics}) literals must be of the datatype {\em xsd:double} whose range should be restricted to be between 0 and 100.
	\begin{itemize}
		\item severity level: ERROR
	\end{itemize}

	\textcolor{blue}{
	\item \textbf{{\em DISCO-C-LITERAL-RANGES-02:}}
{\em disco:cumulativePercentage} (domain: {\em disco:CategoryStatistics}) literals must be of the datatype {\em xsd:double} whose range should be restricted to be between 0 and 100.
	\begin{itemize}
		\item severity level: ERROR
	\end{itemize}
	}

@@

	\item \textbf{{\em DISCO-C-MATHEMATICAL-OPERATIONS-02}}:
	For a given variable, the sum of the frequencies of all codes of the variable's code list has to be equal to the variable's total number of cases (summary statistics of the type 'number of cases').
	\begin{itemize}
		\item severity level: ERROR
	\end{itemize}

	\item \textbf{{\em DISCO-C-MATHEMATICAL-OPERATIONS-03}}:
	For a given variable, the sum of 'valid cases' and 'invalid cases' has to be equal to the total 'number of cases'.
	\begin{itemize}
		\item severity level: ERROR
	\end{itemize}

	\item \textbf{{\em DISCO-C-MATHEMATICAL-OPERATIONS-04}}:
	For a given variable, the total 'number of cases' value for the country 'All' must be equal to the sum of the total 'number of cases' value for each country.
	\begin{itemize}
		\item severity level: ERROR
	\end{itemize}

@@

\begin{itemize}
	\item \textbf{{\em DISCO-C-AGGREGATION-01}}: calculate the number of theoretical concepts in the thematic classification of a given study.
	\begin{itemize}
		\item severity level: INFO
	\end{itemize}
  \item \textbf{{\em DISCO-C-AGGREGATION-02}}: calculate the number of variables of a data set.
	\begin{itemize}
		\item severity level: INFO
	\end{itemize}
	\item \textbf{{\em DISCO-C-AGGREGATION-03}}: calculate the number of questions in a given questionnaire.
	\begin{itemize}
		\item severity level: INFO
	\end{itemize}
	\item \textbf{{\em DISCO-C-AGGREGATION-04}}: the number of codes of a given variable must be below a maximum value.
	\begin{itemize}
		\item severity level: INFO
	\end{itemize}
	\item \textbf{{\em DISCO-C-AGGREGATION-05}}: the number of questions of a given questionnaire must exactly be a given value.
	\begin{itemize}
		\item severity level: INFO
	\end{itemize}
	\item \textbf{{\em DISCO-C-AGGREGATION-06}}: the sum of percentages of all codes of a given variable must be 100.
	\begin{itemize}
		\item severity level: INFO
	\end{itemize}
  \item \textbf{{\em DISCO-C-AGGREGATION-07}}: the absolute frequency of all valid codes of a given variable must be equal to a given value.
	\begin{itemize}
		\item severity level: INFO
	\end{itemize}
\end{itemize}

\section{Data Model Specific RDF Constraint Types}
\label{data-model-specific-constraint-types}

\subsection{Comparison}

\subsection{Data Model Consistency}
Is the data consistent with the intended semantics of the data model?
Such validation rules ensure the integrity of the data according to the data model.

\begin{itemize}
	\item \textbf{{\em DISCO-C-DATA-MODEL-CONSISTENCY-01}}: 
	Codes (\emph{skos:Concept}) are ordered and therefore have fixed positions in an ordered collection (\emph{skos:OrderedCollection}) within a variable representation.
	The cumulative percentage of the current code is the cumulative percentage of the previous code (\emph{disco:cumulativePercentage})
	plus the percentage value (\emph{disco:percentage}) of the current code. 
		\begin{itemize}
		\item severity level: ERROR
	\end{itemize}
	\item \textbf{{\em DISCO-C-DATA-MODEL-CONSISTENCY-02}}: 
	The cumulative percentage (\emph{disco:cumulativePercentage}) of the last code must be 100. 
		\begin{itemize}
		\item severity level: ERROR
	\end{itemize}
	\item \textbf{{\em DISCO-C-DATA-MODEL-CONSISTENCY-03}}: 
	The number of valid cases (\emph{disco:SummaryStatistics} of the type (\emph{disco:summaryStatisticType}) \emph{ddicv-sumstats:ValidCases}) 
	for a particular variable must exactly be the sum of all frequencies of all valid cases (\emph{disco:inValid} of \emph{skos:Concept} is true).
		\begin{itemize}
		\item severity level: ERROR
	\end{itemize}
	\item \textbf{{\em DISCO-C-DATA-MODEL-CONSISTENCY-04}}: 
	The number of invalid cases (\emph{disco:SummaryStatistics} of the type (\emph{disco:summaryStatisticType}) \emph{ddicv-sumstats:InvalidCases})
	for a particular variable must exactly be the sum of all frequencies of all invalid cases (\emph{disco:inValid} of \emph{skos:Concept} is false).
		\begin{itemize}
		\item severity level: ERROR
	\end{itemize}
	\item \textbf{{\em DISCO-C-DATA-MODEL-CONSISTENCY-05}}: 
	The total number of cases (\emph{rdf:value} of the \emph{disco:SummaryStatistics} resource of the type (\emph{disco:summaryStatisticType}) \emph{ddicv-sumstats:NumberOfCases}) 
	for a particular variable must exactly be the number of valid cases plus the number of invalid cases.
		\begin{itemize}
		\item severity level: ERROR
	\end{itemize}
	\item \textbf{{\em DISCO-C-DATA-MODEL-CONSISTENCY-06}}: 
	Some summary statistics types can only be calculated for given variable types. 
	It is not possible to compute minimum values for string variables.
		\begin{itemize}
		\item severity level: ERROR
	\end{itemize}
	\item \textbf{{\em DISCO-C-DATA-MODEL-CONSISTENCY-07}}: 
	Some summary statistics types can only be calculated for given variable types. 
	It is not possible to compute mean values for categorical variables, only for metric variables.
		\begin{itemize}
		\item severity level: ERROR
	\end{itemize}
\end{itemize}

\subsection{Structure}

@@

\subsection{Labeling and Documentation}

@@

\subsection{Vocabulary\textsuperscript{T}}
Vocabularies should not invent any new terms or use deprecated elements.

@@

\subsection{Linked Data}
